% Not to exceed 750 words
\begin{abstract}
The current generation of radial-velocity spectrographs are at the precipice of discovering the first Earth-like exoplanets orbiting in the habitable zones of nearby stars. Such detections require Doppler precision of approximately 10~\si{\centi\meter\per\second}, an order of magnitude better than the typical best-case measurement from the previous generation of instruments. Therefore, the radial-velocity community requires research and innovation from all angles to push our technology over the brink. This thesis presents multiple contributions to this field---ranging from the development of precision laser equipment to the implementation of advanced statistical data analysis algorithms---all in support of the EXtreme PREcision Spectrograph (EXPRES) with the goal of improving instrument precision and exoplanet detection capability.

In Chapter \ref{chapter:modal-noise}, we demonstrate the effectiveness of quasi-chaotic high-amplitude agitation as an optimal form of modal noise mitigation in the optical fibers that feed into radial-velocity spectrographs. This technique is shown to improve radial-velocity error for a single-wavelength laser line from more than 10~\si{\meter\per\second} to less than 60~\si{\centi\meter\per\second} without affecting focal ratio degradation within the fiber. After development of an agitator based on this method for use with EXPRES, we find that combined radial-velocity precision across an entire laser frequency comb improves from 32.8~\si{\centi\meter\per\second} to 6.6~\si{\centi\meter\per\second}.

In Chapter \ref{chapter:astro-comb}, I present aluminum nitride as a nonlinear optical material that can support frequency comb development from near-infrared to ultraviolet wavelengths. By injecting light from an aluminum nitride micro-ring into EXPRES, I demonstrate the material's viability of producing resolvable comb lines throughout the bandpass of the instrument. I also prototype a 16~\si{\giga\hertz} electro-optic modulation comb in combination with an aluminum nitride waveguide as a device that could become a cheap broadband visible-wavelength astro-comb for radial-velocity spectrograph wavelength calibration.

Finally, in Chapters \ref{chapter:pipeline} and \ref{chapter:pipeline2}, I present the EXPRES data extraction pipeline and the numerous novel algorithms that went into its design. Through the default version of the pipeline, including a flat-relative optimal extraction and chunk-by-chunk forward model radial-velocity measurement, we achieve 30~\si{\centi\meter\per\second} single-measurement precision on observations of stars with a signal-to-noise ratio of 250 measured at 550~\si{\nano\meter}. As demonstrated with 51 Peg b, the residual scatter of these observations after fitting with a single-planet Keplerian orbit is less than 90~\si{\centi\meter\per\second}. As alternatives to the default techniques, I also present my implementations of flat-relative spectro-perfectionism and \textit{B}-spline regression stellar template forward modeling within the EXPRES pipeline. These methods provide comparable radial-velocity precision on observations of HD 3651 while also opening up many possibilities for future explorations with radial-velocity data analysis.
\end{abstract}