\begin{center}
{\bf \large Acknowledgments}
\end{center}

I would like to thank Debra, my advisor, for providing me with this opportunity to do such exciting work. I will never forget talking with you the summer before my first year of graduate school, having never touched the field of exoplanets before, and coming away from it absolutely invigorated by the thought of working with you. After almost all of our discussions since then, I leave with that exact same feeling. Thank you for your seemingly endless support, gratitude, and enthusiasm for this wonderful field.

I would also like to thank Andy for being a true mentor for me these past few years. I always appreciated when you stopped by my office for a quick chat, especially since it would, almost inevitably, lead into a one-hour deep dive on some crazy topic one of us had been thinking about. I have truly missed these spontaneous discussions and your corresponding mentorship many times throughout the pandemic.

I would also like to thank members of the extended EXPRES team (both past and present): Allen, Colby, Dave, Joe, Joel, John Michael, Lily, Ren\'e, Ryan B., Sam, and Tyler. Much of the material in this thesis would not have been possible without your guidance and collaboration.

Thank you to the many friends I've made since arriving at Yale six years ago. Chris, Danny, Jeremy, Mariel, Paul, and Will: \textit{you} are what made my first years here even remotely possible and---especially with the addition of Duncan, Hannah, and Kelly---every year has been better than the last. I look forward to seeing all of your future successes as we remain close friends.

Thank you also to my extended New Haven family: Annalisa, Annie, Berk, Cary, Ethan, Jeanette, Kayleigh, Lex, Max, Ratan, Sam, Victor, and many, many others. You all helped me fall in love with this beautiful city and I will always think fondly about this time we spent together here.

I would also like expres my deepest appreciation for my GSA co-leaders/friends Lucy and Maria. You demonstrated that the good faith actions of a few individuals can absolutely make a difference, even in a place as gargantuan as an Ivy League institution.

Finally, Mom, Dad, and Lauren: thank you for being my true constant throughout the entirety of this highly-variable process. Though we may have been states away, I always knew I could count on you to keep me motivated and help me reach the light at the end of the telescope.